---
title: "WUP_04"
author: "David Barton"
date: "fall 2025"
output:
  html_document:
    df_print: paged
---

```{r include=FALSE,echo=FALSE}
require(knitr)
require(tidyverse)
require(tigerstats)
GSScsv.csv <- read_csv("GSScsv.csv")
```

# Introduction

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->

The data is in the file GSScsv.csv and came from the 2018 General Social Survey.  The question to answer is to resolve this hypothesis.  
$H_0$ There is no difference between the average score on the WORDSUM test and level of schooling DEGREE  
$H_a$ There is a significant difference in average score on the WORDSUM test between at least two levels of DEGREE  

# Methods

<!--Mention that you intend to use chatGPT to the greatest extent possible. Give credit to chatGPT when you use it to create code in your report. Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method.!-->

For this assignment I used Numerical Categorical since it was the best equipped for this. 

# Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

## Descriptive Results

### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
basicCN <- ggplot(GSScsv.csv,aes(x=DEGREE,y=WORDSUM))
```
```{r}
basicCN + geom_boxplot()
```


### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
favstats(WORDSUM~DEGREE,data=GSScsv.csv)
```

After looking at favstats. It is clear that Q1 of grad schools is much higher than Q3 of High School and LT High School, meaning there is a difference between them. 

## Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

```{r}
model1 <- aov(WORDSUM~DEGREE,data=GSScsv.csv)
summary.aov(model1)
```

After looking at the inferential results, I saw the P-Vaule is 0.000236, which shows me that there is a different in at least 2 categories


```{r}
par(mar=c(3,17,2,2))
plot(TukeyHSD(model1,conf.level=0.98),las=2,)
```

After plotting a 98% confidence level, i saw the onlt confindence interval that zero does not pass through are LT High School and High School Graduate. This means they are different.

# Discussion and Conclusion

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->

After this assignment. I learned that numerical categorical is able to deal with two different categories. I analyzed the different categories. This was comparing degree to wordsum. I was able to plot the graph. It showed that that the High School for Q1 is higher than Q3 for LT High School and High School Grad. The graphs plotted proved that. Even with the 98% Confidence level.
